---
title: "Overview of farsutil Package"
author: "Richard Sedger"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r global options}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,warning = FALSE)
```

```{r setenv, echo=FALSE, message=FALSE}
library(farsutil)
```

This is a simple package which provides functions to summarize data from the National Highway Traffic Safety Administration (NHTSA) Fatality Analysis Reporting System (FARS) for the years 2013 to 2015. The package was written for the final assignment of the "Building R Packages" course on Coursera, a part of the Johns Hopkins University "Mastering Software Development in R" specialization taught by Roger Peng (Bloomberg School of Public Health at Johns Hopkins University) and Brooke Anderson (Colorado State University).

## The Data

The data in this package come from the National Highway Traffic Safety Administration (NHTSA) Fatality Analysis Reporting System (FARS). It contains the fatal vehicle crashes in the United States for each year. Each observation has as many as 52 features, but the functions in farsutil are concerned only with summarising the total numbers of fatalities by month and State.

Detailed information about the data as a whole can be found in the NHTSA FARS user manuals and documentaion. Please refer to https://crashstats.nhtsa.dot.gov/#/DocumentTypeList/4 


## Summarizing the FARS Data Using the farsutil Package

The `fars_summarize_years()` function takes either a single year, in integer form, or a vector of years, and produces a summary of the simple counts of fatalities by month and year: 

```{r summarize_data}
fars_summarize_years(2013:2014)
```


## Mapping Fatal Crashes by U.S. State

Alternatively, you can use the `fars_map_state` function to produce a map of the fatalities in a particular state. This function takes a state ID number and a year. We need to load the maps package for this function.

For a list of the state ID numbers, see page 26 of the [FARS Analytical User's Guide (2015)](https://crashstats.nhtsa.dot.gov/Api/Public/ViewPublication/812315#27). 

```{r mapping_fatalities}
require(maps)
fars_map_state(48, 2014)  # Texas
fars_map_state(6, 2015)   # California
```

